---
title: "RNA-seq_DNJ4"
author: "Chris"
date: "13/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This script is analysing Linda's DNJ4 mtuant RNASeq data under hydroxyurea conditions of 0mM and 25mM.

Analysis preformed:
Principle Component Analysis
Differential Experssion Analysis 
  

#First, Specify the comparison you would like to look at. 

-Primary-
  Comparison 1: WT vs. DNJ4
  Comparison 2: WTHU vs. DNJ4HU
  Comparison 3: WT vs. WTHU
  Comparison 4: DNJ4 vs. DNJ4HU
  
#Second, Specify any unique gene of which the normalzied expression is of interest to you.
```{r, Input Variabels, include=TRUE, echo=FALSE}
comparison = "4"

```

```{r, loading libraries}
library("tidyverse")
library("gridExtra")
library("ggplot2")
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
    #BiocManager::install("DESeq2")

library('DESeq2')
library("dplyr")
library("rlang")

```


```{r, Reading in the Gene Counts}
#Changing the directory to where the gene counts are
setwd("C:/Users/cwjle/Desktop/GradWork/Data/2020/RNA-Seq/DNJ4_hydroxyurea/ReadsPerGene")

#reading in the list of files in the directory
temp = list.files()
#importing all the counts into a list of dataframes
myfiles = lapply(temp, read.table)

#keeping the sample names of the files
names(myfiles) = temp

#Getting the unstranded counts and removing the first 4 rows of metadata
for (i in 1:length(myfiles)){
  myfiles[[i]] = myfiles[[i]][-c(1:4),c(1:2)]
  colnames(myfiles[[i]]) = c("GeneID",temp[[i]])
}

#Gathering all the counts within a single dataframe
AllCounts = cbind(myfiles[[1]],myfiles[[2]][2],myfiles[[3]][2],myfiles[[4]][2],myfiles[[5]][2],myfiles[[6]][2],myfiles[[7]][2],myfiles[[8]][2],myfiles[[9]][2],myfiles[[10]][2],myfiles[[11]][2],myfiles[[12]][2])

#Changing the gene name column to the row names and getting rid of the gene name column
rownames(AllCounts) = AllCounts$GeneID
AllCounts = AllCounts[,-1]

#writing this dataframe to a counts.csv file. 
setwd("C:/Users/cwjle/Desktop/GradWork/Data/2020/RNA-Seq/DNJ4_hydroxyurea")
write.csv(as.data.frame(AllCounts), file="Allgenecounts.csv")

```
```{r}
#Reading in the metadata
coldata = read.csv("coldata.csv",header = TRUE)

#Moving the samples names from a column to  the rownames
rownames(coldata) = coldata$X
coldata = coldata[,-1]


#building the dds object using the AllCounts, coldata, and strain as a placeholder design factor that we can change later.
dds <- DESeqDataSetFromMatrix(countData = AllCounts,
                              colData = coldata,
                              design = ~ strain)

#Creating a column called group for identification of strain and treatment groups.
dds$group = factor(paste(coldata$strain,coldata$treatment,sep = "_"))

#Changing the design factor to group instead of strain. This way we can pull out the comparisons we want. 
design(dds) = ~group

#getting rid of genes with counts lower than 1
keep <- rowSums(counts(dds)) >= 1
dds <- dds[keep,]

#Running the DESEQ funciton to generate log2fold changes
dds <- DESeq(dds)

```
-Primary-
  Comparison 1: WT vs. DNJ4
  Comparison 2: WTHU vs. DNJ4HU
  Comparison 3: WT vs. WTHU
  Comparison 4: DNJ4 vs. DNJ4HU
  
```{r, Filtering by user input, include=TRUE, echo=FALSE}
#extracting comparison information based on the comparison variable specified at the top of the script. Using contrast function to pull out two groups of interest.
#Specifying comparison in a more descriptive name to make subdirectories for output files and plots.

if  (comparison == 1){
  res <- results(dds, contrast=c("group","DNJ4_control","H99_control"))
  comparison = "WT vs DNJ4"
  
} else if (comparison == 2){
  res <- results(dds, contrast=c("group","DNJ4_25mM_HU","H99_25mM_HU"))
  comparison = "WTHU vs DNJ4HU"
  
} else if (comparison == 3){ 
  res <- results(dds, contrast=c("group","H99_control","H99_25mM_HU"))
  comparison = "WT vs WTHU"
  
} else if (comparison == 4){
  res <- results(dds, contrast=c("group","DNJ4_control","DNJ4_25mM_HU"))
  comparison = "DNJ4 vs DNJ4HU"
  
}

#Creating subfolder and changing working directory to it.
dir.create(paste(comparison))
```

```{r}
setwd(paste(comparison))
#Changing the results object to a data.frame and exporting the raw Log2FoldChange results along with genes with pvalue < 0.05
res = as.data.frame(res)

res_PNA <- as.data.frame(res[,c(2:5)])
write.csv(res_PNA, file=paste(comparison,"_Raw_results.csv"))

#filtering for significant genes.
res_SIG = as.data.frame(res) %>%
  filter(pvalue < 0.05)
write.csv(res_SIG, file=paste(comparison,"_pvalue0.05_results.csv"))


```

